CrypTool Script \\ Mathematics and Cryptography

### 2003-03-10 script/en/chapters/primes.tex
be updates

Although there are only $92$ natural chemical elements, the number of prime
numbers is unlimited. Even the Greek, \index{Euclid} Euclid
\footnote{Euclid,
a Greek mathematician of 4th and 3rd century before Christ. He worked at the
Egyptian academy of Alexandria and wrote ``The Elements'', the most well 
known systematically textbook of the Greek mathematics.} 
knew this in the third century before Christ.
\begin{theorem}[Euclid]\label{thm-pz-euklid}\footnote{The common usage of the term does not denote Euclid as the inventor of the theorem rather;
the true inventor is merely not as prominent. The theorem has already been distinguished
and proven in Euclidís Elements (Book IX, s. 20). The phraseology is remarkable due to 
the fact that the word infinite is not used. The text reads as followed
$$
O\acute{\iota}~\pi\varrho\tilde{\omega}\tau o \iota~\grave{\alpha}\varrho\iota\vartheta\mu o\grave{\iota}~
\pi\lambda\varepsilon\acute{\iota}o \upsilon\varsigma~\varepsilon\grave{\iota}\sigma\grave{\iota}~
\pi\alpha\nu\tau\grave{o}\varsigma~\tau o \tilde{\upsilon}~
\pi\varrho o \tau\varepsilon\vartheta\acute{\varepsilon}\nu\tau o \varsigma~
\pi\lambda\acute{\eta}\vartheta\ o \upsilon\varsigma~
\pi\varrho\acute{\omega}\tau\omega\nu~
\grave{\alpha}\varrho\iota\vartheta\mu\tilde{\omega}\nu,
$$
the English translation of which is: the prime numbers are more than 
any previously existing amount of prime numbers.
} % Ende der Fuﬂnote
The sequence of prime numbers does not discontinue.
Therefore, the quantity of prime numbers is infinite.
\end{theorem}
His proof that there is an infinite number of
primes is still considered to be a brilliant mathematical consideration and
conclusion today (proof by contradiction). He assumed that there is only a
finite number of primes and therefore a largest prime number. Based on this
assumption, he drew logical conclusions until he obtained an obvious
contradiction. This meant that something must be wrong. As there were no
mistakes in the chain of conclusions, it could only be the assumption that was
wrong. Therefore, there must be an infinite number of primes!

@@

These numbers can be found at the following URLs:
\begin{itemize}
  \item[] \href{http://reality.sgi.com/chongo/prime/prime_press.html}
               {\texttt{http://reality.sgi.com/chongo/prime/prime\_press.html}} 
  \item[] \href{http://www.utm.edu/}
               {\texttt{http://www.utm.edu/}}
\end{itemize}

@@

Detailed descriptions of tests for finding out whether a number is prime
can be found on Web sites such as:
\begin{itemize}
  \item[] \href{http://www.utm.edu/research/primes/mersenne.shtml}
               {\texttt{http://www.utm.edu/research/primes/mersenne.shtml}} \\
          \href{http://www.utm.edu/research/primes/prove/index.html}
               {\texttt{http://www.utm.edu/research/primes/prove/index.html}} 
\end{itemize}

@@

Ideas for this can be found at
\begin{itemize}
  \item[] {\href{http://www.utm.edu/research/primes/notes/faq/p_n.html}
          {\tt http://www.utm.edu/research/primes/notes/faq/p\_n.html}}.
\end{itemize}

@@

    More details are in
    \begin{itemize}
      \item[] \href{http://www.fermatsearch.org/history/cosgrave_record.htm/}
          {\texttt{http://www.fermatsearch.org/history/cosgrave\_record.htm/}}
    \end{itemize}

@@

    More details are in
    \begin{itemize}
      \item[] \href{http://www.prothsearch.net/index.html}
                   {\texttt{http://www.prothsearch.net/index.html}} \\
              \href{http://perso.wanadoo.fr/yves.gallot/primes/index.html}
               {\texttt{http://perso.wanadoo.fr/yves.gallot/primes/index.html}} 
    \end{itemize}

@@

Under the heading {\em How Many Primes Are There} at \\
{\href{http://www.utm.edu/research/primes/howmany.shtml}
{\tt http://www.utm.edu/research/primes/howmany.shtml}} \\
you will find numerous other details.

Using the following Web site: \\
{\href{http://www.math.Princeton.EDU/~arbooker/nthprime.html}
{\tt http://www.math.Princeton.EDU/\~{}arbooker/nthprime.html}} \
you can easily determine $PI(x)$.

The \textbf{distribution} of primes displays several irregularities for which no
ÄsystemÄ has yet been found: On the one hand, many occur closely together, like
$2$ and $3,$ $ 11$ and $13, $ $ 809$ and $811$, on the other hand large gaps
containing no primes also occur. For example, no primes lie between $113$ and
$127, $ $ 293$ and $307, $ $317$ and $331, $ $ 523$ and $541, $ $ 773$ and $787,
$ $ 839$ and $853$ as well as between $887$ and $907$. \\
For details, see: \\
{\href{http://www.utm.edu/research/primes/notes/gaps.html}{\tt
http://www.utm.edu/research/primes/notes/gaps.html}}.

@@

Details of this can be found under:

{\href{http://www.cerias.purdue.edu/homes/ssw/cun}{\tt
http://www.cerias.purdue.edu/homes/ssw/cun}}

@@

Twin primes are prime numbers whose difference is 2. Examples include 5 and 7 or
101 and 103. Triplet primes, however, only occur once: 3, 5, 7.  For all other
sets of three consecutive uneven numbers, one of them is always divisible by 3
and thus not a prime.
\begin{itemize}
  \item The number of twin primes is an open question: infinite or limited
number?
  The largest twin primes known today are $1,693,965 \cdot 2^{66,443} \pm 1.$
  \item Does a formula exist for calculating the number of twin primes per
interval?
  \item The above proof of the function  $f: \\qquad N \rightarrow Z$ with $n \mapsto
a^{3^n}$  only guarantees the existence of such a number $a$.  How can we
determine this number $a$ and will it have a value, making the function also of
practical interest?
\item Is there an infinite number of Mersenne prime numbers?\index{Prime number!Mersenne}\index{Mersenne!Mersenne prime number}
\item Is there an infinite number of Fermat prime numbers?

@@

\begin{tabular}{|l|l|l|}\hline
Intervall & Number & Average number per 1000 \\ \hline
1 - 10,000          &    1,229        &    122.900 \\
1 - 100,000         &    9,592        &    95.920 \\
1 - 1,000,000       &    78,498      &    78.498 \\
1 - 10,000,000       &   664,579     &    66.458 \\
1 - 100,000,000      &   5,761,455   &    57.615 \\
1 - 1,000,000,000    &   50,847,534  &    50.848 \\
1 - 10,000,000,000   &   455,052,512 &    45.505 \\ \hline
\end{tabular}
\vskip +6 pt

@@

\bibitem[Bundschuh1998]{Bundschuh1998} \index{Bundschuh 1998}
    Peter Bundschuh, \\
    {\em Einf"uhrung in die Zahlentheorie}, Springer 1988, 4th edition 1998.

>>>>

Although there are only $92$ natural chemical elements, the number of prime
numbers is unlimited. Even the Greek, \index{Euclid} Euclid%
\footnote{Euclid,
a Greek mathematician of 4th and 3rd century before Christ. He worked at the
Egyptian academy of Alexandria and wrote ``The Elements'', the most well 
known systematically textbook of the Greek mathematics.} 
knew this in the third century before Christ.
\begin{theorem}[Euclid\footnote{The common usage of the term does not denote Euclid as the inventor of the theorem rather;
the true inventor is merely not as prominent. The theorem has already been distinguished
and proven in Euclidís Elements (Book IX, s. 20). The phraseology is remarkable due to 
the fact that the word infinite is not used. The text reads as followed
$$
O\acute{\iota}~\pi\varrho\tilde{\omega}\tau o \iota~\grave{\alpha}\varrho\iota\vartheta\mu o\grave{\iota}~
\pi\lambda\varepsilon\acute{\iota}o \upsilon\varsigma~\varepsilon\grave{\iota}\sigma\grave{\iota}~
\pi\alpha\nu\tau\grave{o}\varsigma~\tau o \tilde{\upsilon}~
\pi\varrho o \tau\varepsilon\vartheta\acute{\varepsilon}\nu\tau o \varsigma~
\pi\lambda\acute{\eta}\vartheta\ o \upsilon\varsigma~
\pi\varrho\acute{\omega}\tau\omega\nu~
\grave{\alpha}\varrho\iota\vartheta\mu\tilde{\omega}\nu,
$$
the English translation of which is: the prime numbers are more than 
any previously existing amount of prime numbers.
}]\label{thm-pz-euklid} % Ende der Fuﬂnote
% Fussnote vor ], damit kein Leerraum vor dem Fussnotennummer
The sequence of prime numbers does not discontinue.
Therefore, the quantity of prime numbers is infinite.
\end{theorem}
His proof that there is an infinite number of
primes is still considered to be a brilliant mathematical consideration and
conclusion today (proof by contradiction). He assumed that there is only a
finite number of primes and therefore a largest prime number. Based on this
assumption, he drew logical conclusions until he obtained an obvious
contradiction. This meant that something must be wrong. As there were no
mistakes in the chain of conclusions, it could only be the assumption that was
wrong. Therefore, there must be an infinite number of primes!

@@

These numbers can be found at the following URLs:
\vspace{-10pt}
\begin{itemize}
  \item[] \href{http://reality.sgi.com/chongo/prime/prime_press.html}
               {\texttt{http://reality.sgi.com/chongo/prime/prime\_press.html}}\\
          \href{http://www.utm.edu/}
               {\texttt{http://www.utm.edu/}}
\end{itemize}

@@

Detailed descriptions of tests for finding out whether a number is prime
can be found on Web sites such as:
\vspace{-10pt}
\begin{itemize}
  \item[] \href{http://www.utm.edu/research/primes/mersenne.shtml}
               {\texttt{http://www.utm.edu/research/primes/mersenne.shtml}} \\
          \href{http://www.utm.edu/research/primes/prove/index.html}
               {\texttt{http://www.utm.edu/research/primes/prove/index.html}} 
\end{itemize}

@@

Ideas for this can be found at
\vspace{-10pt}
\begin{itemize}
  \item[] {\href{http://www.utm.edu/research/primes/notes/faq/p_n.html}
          {\tt http://www.utm.edu/research/primes/notes/faq/p\_n.html}}.
\end{itemize}

@@

    More details are in
    \vspace{-10pt}
    \begin{itemize}
      \item[] \href{http://www.fermatsearch.org/history/cosgrave_record.htm/}
          {\texttt{http://www.fermatsearch.org/history/cosgrave\_record.htm/}}
    \end{itemize}

@@

    More details are in
    \vspace{-10pt}
    \begin{itemize}
      \item[] \href{http://www.prothsearch.net/index.html}
                   {\texttt{http://www.prothsearch.net/index.html}} \\
              \href{http://perso.wanadoo.fr/yves.gallot/primes/index.html}
               {\texttt{http://perso.wanadoo.fr/yves.gallot/primes/index.html}} 
    \end{itemize}

@@

Under the heading {\em How Many Primes Are There} at
\vspace{-10pt}
\begin{itemize}
  \item[] \href{http://www.utm.edu/research/primes/howmany.shtml}
               {\tt http://www.utm.edu/research/primes/howmany.shtml}
\end{itemize}
you will find numerous other details.

Using the following Web site:
\vspace{-10pt}
\begin{itemize}
  \item[] \href{http://www.math.Princeton.EDU/~arbooker/nthprime.html}
               {\tt http://www.math.Princeton.EDU/\~{}arbooker/nthprime.html}
\end{itemize}

you can easily determine $PI(x)$.

The \textbf{distribution} of primes displays several irregularities for which no
ÄsystemÄ has yet been found: On the one hand, many occur closely together, like
$2$ and $3,$ $ 11$ and $13, $ $ 809$ and $811$, on the other hand large gaps
containing no primes also occur. For example, no primes lie between $113$ and
$127, $ $ 293$ and $307, $ $317$ and $331, $ $ 523$ and $541, $ $ 773$ and $787,
$ $ 839$ and $853$ as well as between $887$ and $907$. \\
For details, see: 
\vspace{-10pt}
\begin{itemize}
  \item[] \href{http://www.utm.edu/research/primes/notes/gaps.html}
               {\tt http://www.utm.edu/research/primes/notes/gaps.html}
\end{itemize}

@@

Details of this can be found under:
\vspace{-10pt}
\begin{itemize}
  \item[] \href{http://www.cerias.purdue.edu/homes/ssw/cun}
               {\tt http://www.cerias.purdue.edu/homes/ssw/cun}
\end{itemize}

@@

Twin primes are prime numbers whose difference is 2. Examples include 5 and 7 or
101 and 103. Triplet primes, however, only occur once: 3, 5, 7.  For all other
sets of three consecutive uneven numbers, one of them is always divisible by 3
and thus not a prime.
\begin{itemize}
  \item The number of twin primes is an open question: infinite or limited
number?
  The largest twin primes known today are $1,693,965 \cdot 2^{66,443} \pm 1.$
  \item Does a formula exist for calculating the number of twin primes per
interval?
  \item The above proof of the function  $f: \: N \rightarrow Z$ with $n \mapsto
a^{3^n}$  only guarantees the existence of such a number $a$.  How can we
determine this number $a$ and will it have a value, making the function also of
practical interest?
\item Is there an infinite number of Mersenne prime numbers?\index{Prime number!Mersenne}\index{Mersenne!Mersenne prime number}
\item Is there an infinite number of Fermat prime numbers?

@@

\begin{tabular}{|l|l|l|}\hline
Interval & Number & Average number per 1000 \\ \hline
1 - 10,000          &    1,229        &    122.900 \\
1 - 100,000         &    9,592        &    95.920 \\
1 - 1,000,000       &    78,498      &    78.498 \\
1 - 10,000,000       &   664,579     &    66.458 \\
1 - 100,000,000      &   5,761,455   &    57.615 \\
1 - 1,000,000,000    &   50,847,534  &    50.848 \\
1 - 10,000,000,000   &   455,052,512 &    45.505 \\ \hline
\end{tabular}
\vskip +6 pt

@@

\bibitem[Bundschuh1998]{Bundschuh1998} \index{Bundschuh 1998}
    Peter Bundschuh, \\
    {\em Einf\"uhrung in die Zahlentheorie}, Springer 1988, 4th edition 1998.
